EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-23 02_install.tex
Intermediate GB almost finished with the KMH example LC doing something here and there, as I see

\documentclass[edipack2.tex]{subfiles}
\begin{document}

\section{Installation}\label{SecInstall}
The configuration and installation of \NAME is handled by CMake, which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is a static Fortran library called {\tt  libedipack.a} which, alongside the compiled
modules, wraps the \NAME software in both its single- and multi-impurity variants.
A second dynamic library, {\tt libedipack\_cbindings.so}, together with the associated header
file{\tt edipack\_cbindings.h}, enables interoperability with other programming languages.

@@

\begin{itemize}
\item{\bf EDIpack.}
  This module forms the foundation of the library, implementing the 
  Lanczos-based solver for general quantum impurity problems. It 
  supports systems with a wide range of symmetries, including quantum 
  number conservation, multi-orbital models, and even coupling to local 
  phonons. The \NAME solver is structured hierarchically and modularly, 
  with different sections of the library communicating through a shared 
  memory layer. The top-level module, {\tt EDIPACK}, provides access 
  to the core Fortran API, exposing key procedures for initialization, 
  execution, and finalization, while abstracting the underlying data 
  structures. A detailed overview of this part of the library is 
  provided in Sec.\ref{SecEDIpack}.

>>>>

\documentclass[edipack_sp.tex]{subfiles}
\begin{document}

\section{Installation}\label{SecInstall}
The configuration and installation of \NAME is handled by CMake, which ensures
multi-platforms compabitility and dependencies resolution.  
The software builds into two distinct libraries.
The main one is a static Fortran library called {\tt  libedipack.a} which, alongside the compiled
modules, wraps the \NAME software in both its single- and multi-impurity variants.
A second dynamic library, {\tt libedipack\_cbindings.so}, together with the associated header
file {\tt edipack\_cbindings.h}, enables interoperability with other programming languages.

@@

\begin{itemize}
\item{\bf EDIpack.}
  This module forms the foundation of the library, implementing the 
  Lanczos-based solver for general quantum impurity problems. It 
  supports systems with a wide range of symmetries leveraging different quantum numbers conservation, multi-orbital models, and even coupling to local 
  phonons. The \NAME solver is structured hierarchically and modularly, 
  with different sections of the library communicating through a shared 
  memory layer. The top-level module, {\tt EDIPACK}, provides access 
  to the core Fortran API, exposing key procedures for initialization, 
  execution, and finalization, while abstracting the underlying data 
  structures. A detailed overview of this part of the library is 
  provided in Sec.\ref{SecEDIpack}.

### 2025-05-23 04_cbinding.tex
Intermediate GB almost finished with the KMH example LC doing something here and there, as I see

\subsection{EDIpack2py, the Python API}\label{sSecInteropEDIpy}
As a first application of the \NAME C-bindings we implemented a
complete Python interface, i.e. EDIpack2py. This is a Python module which
enables access to all the library features and unlock implementation of
further interfaces of \NAME as a plug-in solver for external Python
based software.

@@

\paragraph{Anaconda}
As for \NAME, the Python API in EDIpack2py is available as an
Anaconda package for GNU/Linux and Mac Os systems. Packages are available for
Python version $>3.10$. The \NAME package contains the {\tt
  EDIpack2py} Python module as well as \NAME and SciFortran
libraries. In this case the resolution of the dependencies is
taken care from Conda itself. 
Using conda the installation reads:
\begin{lstlisting}[style=mybash]
conda create -n edipack
conda activate edipack
conda install -c conda-forge -c edipack edipack
\end{lstlisting}

Once installed, any loading of the {\tt EDIpack2py} module resolves in the
attempt to load the dynamic library {\tt
  libedipack\_cbindings.so/.dylib} containing the Fortran-C bindings
for \NAME. By default the library is searched proceeds as follow: 
\begin{enumerate}
\item First the user can override the location of the library
  (determined during \NAME build configuration) by exporting an
  environment variable called {\tt EDIPACK\_PATH} .
\item By default, the Python module detects the position of the
  Fortran libraries via {\tt pkg-config}. Any of the OS loading method
  outlined in \secu{sSecInstallOSloading} automatically push the
  correct configuration to the {\tt PKG\_CONFIG\_PATH}. 
\item As a last resort the environment variables {\tt
    LD\_LIBRARY\_PATH} and {\tt DYLD\_LIBRARY\_PATH} are analyzed to
  retrieve the correct position. 
\end{enumerate}
If none of the previous attempts succeeds, the module will not not load correctly and an error message will be printed.

\subsubsection{Implementation}\label{sSecInteropEDIpyImplementation}
The Python API provided in the EDIpack2py module consists essentially of a
suitable class called for convenience {\tt global\_env}.
The class incorporates all the global variables inherited from \NAME
C-bindings library and implements a number of interface functions
leveraging the Python duck typing to \NAME.  
The variables and the functions of \NAME are exposed to the user and
are accessed as properties and
methods of the {\tt global\_env} claass, which needs to be imported at the beginning
of the python script, along with other useful modules. For instance Numpy is
necessary, while mpi4py is strongly recommended.

@@

The EDIpack2py supports the solution of problems with independent
impurities, interfacing with the the EDIpack2ineq extension of the
library, if present. Should the inequivalent impurities package not be
built, the Python module silently disable the support to it, so that
invoking any related procedure  will result in a {\tt RuntimeError}.
The user can check the availability of the inequivalent impurities
interface by querying the value of  {\tt edipack2py.global\_env.has\_ineq}.

@@

\paragraph{{\bf Global variables}}
This includes a subset of the input variables available in \NAME which
are used to control the calculation.
The variables are loaded globally in {\tt EDIpack2py} and can be accessed
or set locally as properties of the class {\tt
  global\_env}. Differently these are initialized, alongside the remaining
default input variables, through a call to procedure {\tt
  edipack2py.global\_env.read\_input} which interface the {\tt
  ed\_read\_input} function in \NAME. 
A given example is reported in the following code extract:

\begin{lstlisting}[language=python,  frame=lines]
import numpy as np
from edipack2py import global_env as ed
ed.Nspin = 1            # set a global variable
mylocalvar = ed.Nspin   # assign to a local variable
print(ed.Nspin)         # all functions can have global variables as arguments
np.arange(ed.Nspin)
\end{lstlisting}

\paragraph{{\bf Solver functions}}
This group includes a number of functions enabling initialization,
execution and finalization  functionalities for the diagonalization
solver.
\begin{itemize}
  \item {\tt init\_solver} and {\tt set\_Hloc}. The first 
    initializes the \NAME environment for the quantum impurity problem
    solution, sets the effective bath either reading it from a file or
    initializing it from flat band model. Once this function is
    called, it is not possible to allocate a second instance of the solver.
    {\tt set\_Hloc} sets the
    non-interacting part of the impurity Hamiltonian $\h^0_{\a\b\s\s'}$. 
    Either function take different argument combinations there
    including support for inequivalent impurities.

  \item {\tt solve} This function solves the quantum impurity problem,
    calculates the observables and any dynamical correlation
    function. All results remain stored in the memory and get accessed
    through input/output functions.

@@

\paragraph{{\bf Constructor}}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode=normal,superc,nonsu2} (Sec.~\ref{sSecQNs}) and
{\tt bath\_type=normal,hybrid,general} (Sec.~\ref{sSecBath}). Besides, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

@@

In order to use the \NAME impurity solver for a calculation instead of the built-in implementation of the CT-HYB algorithm, the user needs to set the {\tt solver} parameter in section {\tt General} to \NAME. Configuration specific to the exact diagonalization solver can be performed through the parameters in section \NAME of a w2dynamics parameter file. For most of the \NAME input variables, it provides a parameter of the same name (in all uppercase letters) that allows the corresponding variable to be set insofar as this is supported by the interface and not an option available through generic w2dynamics parameters. This thus includes some parameters that the user will ordinarily want to change from their defaults in a typical calculation such as the number of bath sites {\tt NBATH} and other parameters such as options for the bath fitting and diagonalization algorithms, but also specifically excludes other typically set input variables such as {\tt NORB}, {\tt ULOC}, {\tt BETA}, and {\tt   NLOOP}. All of the these variables correspond to or follow from configuration parameters that a user also needs to set for a CT-HYB calculation, and in order to present a consistent and rather ``solver-agnostic'' user interface, no redundant parameters are added to the {\tt EDIPACK} section and the usual parameters such as {\tt   beta} in section {\tt General} must be used instead. Note however in particular that if the \NAME solver is used, {\tt beta} is by default only used as cut-off and does not imply a finite temperature calculation unless the user explicitly sets the parameter {\tt   ED\_FINITE\_TEMP} in section {\tt EDIPACK} to {\tt True}.

Solver specificities are abstracted away to the greatest reasonable extent by having the input, output, and DMFT loop handled by w2dynamics in mostly the same way as it would be for the CT-HYB solver. This e.g. means that results such as the Green's functions are stored into the w2dynamics HDF5 output file in the typical format, and that w2dynamics uses only its own code for e.g. calculations with multiple inequivalent impurities without relying on any related implementation provided by \NAME{}. As a consequence of this, some functionality that would require specific support from w2dynamics that is not available as of the time of publication can not be used through the interface, e.g. calculations with superconductivity or with phonons.

>>>>

\subsection{EDIpack2py, the Python API}\label{sSecInteropEDIpy}
As a first application of the \NAME C-bindings we implemented a
complete Python interface called EDIpack2py. This is a Python module which
enables access to all the library features and unlock implementation of
further interfaces of \NAME as a plug-in solver for external Python
based software. Detailed documentation can be found online at \href{https://edipack.github.io/EDIpack2py}
{https://edipack.github.io/EDIpack2py}.

@@

\paragraph{Anaconda}
As for \NAME, the Python API in EDIpack2py is available as an
Anaconda package for GNU/Linux and Mac Os systems. Packages are available for
Python version $>3.10$. The \NAME package contains the {\tt
  EDIpack2py} Python module as well as \NAME and SciFortran
libraries. 
%In this case the resolution of the dependencies is
%taken care from Conda itself. 
Using conda or mamba the installation proceeds as following:
\begin{lstlisting}[style=mybash]
conda create -n edipack  #creation of a virtual environment (name can be changed)
conda activate edipack   #activation of the virtual environment
conda install -c conda-forge -c edipack edipack #installation from the edipack channel
\end{lstlisting}

When the{\tt EDIpack2py} module is imported, it attempts to load the dynamic library {\tt
  libedipack\_cbindings.so/.dylib} containing the Fortran-C bindings
for \NAME. By default the library search proceeds as follow: 
\begin{enumerate}
\item First the user can override the location of the library
  (determined during \NAME build configuration) by exporting an
  environment variable called {\tt EDIPACK\_PATH} .
\item By default, the Python module detects the position of the
  Fortran libraries via {\tt pkg-config}. Any of the OS loading method
  outlined in \secu{sSecInstallOSloading} automatically push the
  correct configuration to the {\tt PKG\_CONFIG\_PATH}. 
\item As a last resort the environment variables {\tt
    LD\_LIBRARY\_PATH} and {\tt DYLD\_LIBRARY\_PATH} are analyzed to
  retrieve the correct position. 
\end{enumerate}
If none of the previous attempts succeeds, the module will not not load correctly and an error message will be printed.

\subsubsection{Implementation}\label{sSecInteropEDIpyImplementation}
The Python API provided in the EDIpack2py module consists essentially of a
class called for convenience {\tt global\_env}.
This class contains all the global variables inherited from the \NAME
C-bindings library, and implements a number of interface functions
leveraging the Python duck typing to \NAME.  
The variables and the functions of \NAME are exposed to the user and
are accessed as properties and
methods of the {\tt global\_env} class.

The {\tt global\_env} class needs to be imported at the beginning
of the python script, along with other useful modules. Numpy is
necessary, while mpi4py is strongly recommended.

@@

The EDIpack2py module supports the solution of problems with independent
impurities, interfacing to the EDIpack2ineq extension of the
library, if present. Should the inequivalent impurities package not be
built, the Python module silently disable the support to it, so that
invoking any related procedure  will result in a {\tt RuntimeError}.
The user can check the availability of the inequivalent impurities
interface by querying the value of  {\tt edipack2py.global\_env.has\_ineq}.

@@

\paragraph{{\bf Global variables}}
This includes a subset of the input variables available in \NAME which
are used to control the calculation.
The variables are loaded globally in {\tt EDIpack2py} and can be accessed
or set locally as properties of the class {\tt
  global\_env}. The global variables are initialized, alongside the remaining
default input variables, through a call to procedure {\tt
  edipack2py.global\_env.read\_input} which interface the {\tt
  ed\_read\_input} function in \NAME. 
A given example is reported in the following code extract:

\begin{lstlisting}[language=python,  frame=lines]
import numpy as np
from edipack2py import global_env as ed
ed.Nspin = 1            # set a global variable
mylocalvar = ed.Nspin   # assign to a local variable
print(ed.Nspin)         # all functions can have global variables as arguments
np.arange(ed.Nspin)     # array of integers from 0 to Nspin-1
\end{lstlisting}

\paragraph{{\bf Solver functions}}
This group includes a number of functions enabling initialization,
execution and finalization  of the \NAME solver.
\begin{itemize}
  \item {\tt init\_solver} and {\tt set\_Hloc}. The first 
    initializes the \NAME environment for the quantum impurity problem
    solution, sets the effective bath either reading it from a file or
    initializing it from flat band model. Once this function is
    called, it is not possible to allocate a second instance of the solver.
    {\tt set\_Hloc} sets the
    non-interacting part of the impurity Hamiltonian $\h^0_{\a\b\s\s'}$. 
    Either function take different argument combinations there
    including support for inequivalent impurities.

  \item {\tt solve} This function solves the quantum impurity problem,
    calculates the observables and any dynamical correlation
    function. All results remain stored in the memory and can be accessed
    through input/output functions.

@@

\paragraph{{\bf Constructor}}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode=normal,superc,nonsu2} (Sec.~\ref{sSecQNs}) and
{\tt bath\_type=normal,hybrid,general} (Sec.~\ref{sSecBath}). In addition, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

@@

In order to use the \NAME impurity solver for a calculation instead of the built-in implementation of the CT-HYB algorithm, the user needs to set the {\tt solver} parameter in section {\tt General} to \NAME.

The  behavior of the w2dynamics solver is controlled by means of an input file, by default called
{\tt Parameters.in}. In order to configure the ED solver, the content of a specific section in the parameters file, called {\tt [EDIPACK]}, has to be modified.

For most of the \NAME input variables, w2dynamics accepts a parameter of the same name (in all uppercase letters) that allows the corresponding variable to be set, and is declared in the {\tt EDIPACK} section of the parameters file. This set of variables includes most of the control parameters to be tuned for a specific DMFT-ED calculation, such as the number of bath sites {\tt NBATH} and options for the bath fitting and diagonalization algorithms.

A second set of input variables, including phyisical quantities such as {\tt NORB}, {\tt ULOC}, {\tt BETA}, and {\tt   NLOOP}, directly map to parameters available in other sections of the w2dynamics input file, such as {\tt [General]} or {\tt [Atoms]}. An example of a properly formatted input file is provided in~\ref{SecExamples}. See~\cite{Wallerberger2019CPC}
for further details.

%All of the these variables correspond to or follow from configuration parameters that a user also needs to set for a CT-HYB calculation, and in order to present a consistent and rather ``solver-agnostic'' user interface, no redundant parameters are added to the {\tt EDIPACK} section and the usual parameters such as {\tt   beta} in section {\tt General} must be used instead. Note however in particular that if the \NAME solver is used, {\tt beta} is by default only used as cut-off and does not imply a finite temperature calculation unless the user explicitly sets the parameter {\tt   ED\_FINITE\_TEMP} in section {\tt EDIPACK} to {\tt True}.

Solver specificities are abstracted away to the greatest reasonable extent by having the input, output, and DMFT loop handled by w2dynamics in a way closely resembling the standard the CT-HYB code. For example, results such as the Green's functions are stored into the w2dynamics HDF5 output file in the typical format, and that w2dynamics uses only its own logic for calculations with multiple inequivalent impurities without relying on any related implementation provided by \NAME{}. As a consequence of this, some functionality that would require specific support from w2dynamics that is not available as of the time of publication can not be used through the interface, for instance calculations with {\tt ED\_MODE=SUPERC} or with phonons.

### 2025-05-23 05_examples.tex
Intermediate GB almost finished with the KMH example LC doing something here and there, as I see

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
  !Setup bath and solver for anisotropic AFM phases
  !> define the symmetry-basis:
  Nineq = 2
  Nsym = 2
  allocate(Hsym_basis(Nspin*Norb,Nspin*Norb,Nsym))
  Hsym_basis(:,:,1) = pauli_0
  select case(ansatz)
    case('off-plane')
        Hsym_basis(:,:,2) = pauli_z
    case('in-plane')
        Hsym_basis(:,:,2) = pauli_x
  end select
  !> initialize the symmetry parameters
  allocate(lambdasym_vectors(Nineq,Nbath,Nsym))
  call build_replica_band(onsite_band,ed_hw_bath,Nbath)
  lambdasym_vectors(:,:,1) = onsite_band
  lambdasym_vectors(:,:,2) = 0d0 ! paramagnetic
  !> define a symmetry-breaking AFM kick, to help numerics
  if(afmkick)then
     lambdasym_vectors(1,:,2) = +sb_field
     lambdasym_vectors(2,:,2) = -sb_field
  endif
  !> setup H_bath inside the solver
  select case(bath_type)
    case('replica')
        call ed_set_Hreplica(Hsym_basis,lambdasym_vectors)
    case('general')
        call ed_set_Hgeneral(Hsym_basis,lambdasym_vectors)
  end select
  !> finally initialize the two solver instances, one per sublattice
  !  (the EDIpack2ineq layer automatically dispatches it for the two
  !   impurity models, if we feed a Bath with the [Nineq,Nb] shape!)
  Nb = ed_get_bath_dimension(Hsym_basis)
  allocate(Bath(Nineq,Nb))
  call ed_init_solver(Bath)
  .
  .
  .
  contains
  .
  .
  .
  subroutine build_replica_band(epsilons,bandwidth,Nreplica)
  !> Initialization of the (average) bath levels, suitable to
  !  describe a weakly interacting metal or a band insulator
      real(8),allocatable,dimension(:,:),intent(out)   :: epsilons
      real(8),intent(in)                               :: bandwidth
      integer,intent(in)                               :: Nreplica
      integer                                          :: ireplica
      real(8),dimension(Nreplica)                      :: tempvec
      real(8)                                          :: tempval
      !
      allocate(epsilons(Nineq,Nbath))
      !
      do ireplica=1,Nreplica
         tempval = ireplica - 1 - (Nreplica-1)/2d0       ![-(N-1)/2 : (N-1)/2]
         tempval = tempval * 2 * bandwidth/(Nreplica-1)  !-bandwidth:bandwidth
         tempvec(ireplica) = tempval
      enddo
      !
      if(mod(Nreplica,2)==0)then
         tempvec(Nreplica/2) = -1d-1  !Much needed small energies around the
         tempvec(Nreplica/2+1) = 1d-1 !Fermi level (if EF itself not present)
      endif
      !
      do ilat=1,Nineq
         epsilons(ilat,:) = tempvec
      enddo
      !
   end subroutine build_replica_band
\end{lstlisting}

@@

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
      iloop=iloop+1
      !
      !> Solve the two inequivalent impurity problems
      if(neelsym)then
        !> solve just one sublattice and get the other by Neel symmetry
        call ed_solve(comm,Bath(1,:),Hloc(1,:,:,:,:))
        call ed_get_sigma_matsubara(Smats(1,:,:,:,:,:))
        select case(ansatz)
            case('off-plane')
             Smats(2,2,2,:,:,:) = Smats(1,1,1,:,:,:) 
             Smats(2,1,1,:,:,:) = Smats(1,2,2,:,:,:) 
             if(master)write(*,*) ">>> Enforcing AFMz symmetry"
            case('in-plane')
             Smats(2,1,1,:,:,:) = Smats(1,1,1,:,:,:)   
             Smats(2,2,2,:,:,:) = Smats(1,2,2,:,:,:)   
             Smats(2,1,2,:,:,:) = -Smats(1,1,2,:,:,:)  
             Smats(2,2,1,:,:,:) = -Smats(1,2,1,:,:,:) 
             if(master)write(*,*) ">>> Enforcing AFMx symmetry"
        end select
      else
         !> solve both sublattices independently with the RDMFT wrapper:
         !  mpi_lanc=T => MPI lanczos, mpi_lanc=F => MPI for ineq sites
         call ed_solve(Bath,Hloc,mpi_lanc=.true.)
         !> retrieve all self-energies:
         call ed_get_sigma_matsubara(Smats,Nlat)
         !
      endif
      !
      !> Compute the local Green's function
      call dmft_gloc_matsubara(Hk,Gmats,Smats)
      !
      !> Compute the two inequivalent Weiss fields
      call dmft_self_consistency(Gmats,Smats,Weiss)
      !
      !> Fit the new bath
      !  this behaves differently depending on the ed_mode input:
      !  - for the normal mode: normal/AFMz phase is solved, so we 
      !    can fit spin1 and spin2 components independently
      !  - for the nonsu2 mode: Sz-conservation can be broken, so a 
      !    unique fit for both spin components is needed
      select case(ed_mode)
       case("normal")
         call ed_chi2_fitgf(Bath,Weiss,ispin=1)
         call ed_chi2_fitgf(Bath,Weiss,ispin=2)
       case("nonsu2")
         call ed_chi2_fitgf(Bath,Weiss,Hloc)
      end select
      !
      !> linear mixing (aids convergence near transition points)
      if(iloop>1) Bath = wmixing*Bath + (1.d0-wmixing)*Bath_prev
      Bath_prev = Bath
      !
      !> check convergence
      converged = check_convergence(Weiss(:,1,1,1,1,:),dmft_error,nsuccess,nloop)
      !
   enddo
\end{lstlisting}

After reaching convergence, we can

>>>>

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
program dmft_kane_mele_hubbard
  !Load EDIpack library 
  USE EDIpack
  !Load the inequivalent impurities extension
  USE EDIpack2ineq
  .
  .
  .
  !Setup bath and solver for anisotropic AFM phases
  !> define the symmetry-basis:
  Nineq = 2
  Nsym = 2
  allocate(Hsym_basis(Nspin*Norb,Nspin*Norb,Nsym))
  Hsym_basis(:,:,1) = pauli_0
  select case(ansatz)
    case('off-plane')
        Hsym_basis(:,:,2) = pauli_z
    case('in-plane')
        Hsym_basis(:,:,2) = pauli_x
  end select
  !> initialize the symmetry parameters
  allocate(lambdasym_vectors(Nineq,Nbath,Nsym))
  call build_replica_band(onsite_band,ed_hw_bath,Nbath)
  lambdasym_vectors(:,:,1) = onsite_band
  lambdasym_vectors(:,:,2) = 0d0 ! paramagnetic
  !> define a symmetry-breaking AFM kick, to help numerics
  if(afmkick)then
     lambdasym_vectors(1,:,2) = +sb_field
     lambdasym_vectors(2,:,2) = -sb_field
  endif
  !> setup H_bath inside the solver
  select case(bath_type)
    case('replica')
        call ed_set_Hreplica(Hsym_basis,lambdasym_vectors)
    case('general')
        call ed_set_Hgeneral(Hsym_basis,lambdasym_vectors)
  end select
  !> finally initialize the two solver instances, one per sublattice
  !  (the EDIpack2ineq layer automatically dispatches it for the two
  !   impurity models, if we feed a Bath with the [Nineq,Nb] shape!)
  Nb = ed_get_bath_dimension(Hsym_basis)
  allocate(Bath(Nineq,Nb))
  call ed_init_solver(Bath)
  .
  .
  .
  contains
  .
  .
  .
  subroutine build_replica_band(epsilons,bandwidth,Nreplica)
  !> Initialization of the (average) bath levels, suitable to
  !  describe a weakly interacting metal or a band insulator
      real(8),allocatable,dimension(:,:),intent(out)   :: epsilons
      real(8),intent(in)                               :: bandwidth
      integer,intent(in)                               :: Nreplica
      integer                                          :: ireplica
      real(8),dimension(Nreplica)                      :: tempvec
      real(8)                                          :: tempval
      !
      allocate(epsilons(Nineq,Nbath))
      !
      do ireplica=1,Nreplica
         tempval = ireplica - 1 - (Nreplica-1)/2d0       ![-(N-1)/2 : (N-1)/2]
         tempval = tempval * 2 * bandwidth/(Nreplica-1)  !-bandwidth:bandwidth
         tempvec(ireplica) = tempval
      enddo
      !
      if(mod(Nreplica,2)==0)then
         tempvec(Nreplica/2) = -1d-1  !Much needed small energies around the
         tempvec(Nreplica/2+1) = 1d-1 !Fermi level (if EF itself not present)
      endif
      !
      do ilat=1,Nineq
         epsilons(ilat,:) = tempvec
      enddo
      !
   end subroutine build_replica_band
\end{lstlisting}

@@

\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
      iloop=iloop+1
      !
      !> Solve the two inequivalent impurity problems
      if(neelsym)then
        !> solve just one sublattice and get the other by Neel symmetry
        call ed_solve(Bath(1,:))
        call ed_get_sigma_matsubara(Smats(1,:,:,:,:,:))
        select case(ansatz)
            case('off-plane')
             Smats(2,2,2,:,:,:) = Smats(1,1,1,:,:,:) 
             Smats(2,1,1,:,:,:) = Smats(1,2,2,:,:,:) 
             if(master)write(*,*) ">>> Enforcing AFMz symmetry"
            case('in-plane')
             Smats(2,1,1,:,:,:) = Smats(1,1,1,:,:,:)   
             Smats(2,2,2,:,:,:) = Smats(1,2,2,:,:,:)   
             Smats(2,1,2,:,:,:) = -Smats(1,1,2,:,:,:)  
             Smats(2,2,1,:,:,:) = -Smats(1,2,1,:,:,:) 
             if(master)write(*,*) ">>> Enforcing AFMx symmetry"
        end select
      else
         !> solve both sublattices independently with the RDMFT wrapper:
         !  mpi_lanc=T => MPI lanczos, mpi_lanc=F => MPI for ineq sites
         call ed_solve(Bath,mpi_lanc=.true.)
         !> retrieve all self-energies:
         call ed_get_sigma_matsubara(Smats,Nineq)
         !
      endif
      !
      !> Compute the local Green's function
      call dmft_gloc_matsubara(Hk,Gmats,Smats)
      !
      !> Compute the two inequivalent Weiss fields
      call dmft_self_consistency(Gmats,Smats,Weiss)
      !
      !> Fit the new bath
      !  this behaves differently depending on the ed_mode input:
      !  - for the normal mode: normal/AFMz phase is solved, so we 
      !    can fit spin1 and spin2 components independently
      !  - for the nonsu2 mode: Sz-conservation can be broken, so a 
      !    unique fit for both spin components is needed
      select case(ed_mode)
       case("normal")
         call ed_chi2_fitgf(Bath,Weiss,ispin=1)
         call ed_chi2_fitgf(Bath,Weiss,ispin=2)
       case("nonsu2")
         call ed_chi2_fitgf(Bath,Weiss,Hloc)
      end select
      !
      !> linear mixing (aids convergence near transition points)
      if(iloop>1) Bath = wmixing*Bath + (1.d0-wmixing)*Bath_prev
      Bath_prev = Bath
      !
      !> check convergence
      converged = check_convergence(Weiss(:,1,1,1,1,:),dmft_error,nsuccess,nloop)
      !
   enddo
   !
   !> Compute Kinetic Energy
   call dmft_kinetic_energy(Hk,Smats)
\end{lstlisting}

We observe that, after convergence is reached, we can perform some 
post-processing to get quantities of interest, e.g. computing the 
lattice kinetic energy (as implemented in the \texttt{dmft\_tools} 
library). The potential energy, as well as the AFM order parameters
and most impurity observables are automatically computed by \NAME, 
and saved with an appropriate inequivalent-impurity filename suffix 
by the EDIpack2ineq layer.

Hence, we can readily evaluate the total energy of both the AFM$_\perp$
and AFM$_\parallel$ calculations and compare them to establish which 
is favored as the groundstate of the system. In \figu{fig:KMenergy}(a)
we show results for the energy difference $E_\parallel-E_\perp$ across
a wide range of $U/t$ and $\lambda_\mathrm{so}$ values. To assess the
degree of accuracy of our total energy calculation, we benchmark against
Hartree-Fock (simple static mean-field, but inherently variational), 
DMFT with NRG solver and lattice AFQMC data, as taken from 
Ref.\,\cite{Racz2020}. The total energy comparison is reported in 
\figu{fig:KMenergy}(b), for $\lambda_\mathrm{so}=0$.

\begin{figure}
\hspace{1cm} (a) \hspace{7.25cm} (b)\\
    \includegraphics[width=0.47\linewidth,trim={0 0 0 5mm},clip]{figures/KMH_energy.pdf}\hfill
    \includegraphics[width=0.45\linewidth]{figures/benchEnergy_honey.pdf}\\
    \caption{In panel (a) the total energy difference between the in-plane 
    and off-plane AFM solutions found in DMFT-ED at $T=0$. Negative values
    signal a groundstate with in-plane magnetization. The plateau at zero
    energy difference marks the topological phase of the model (a 
    paramagnetic quantum spin-Hall insulator), at all points except on the $\lambda_\mathrm{so}=0$ line, where the groundstate is an interacting
    Dirac liquid, at weak coupling, or an isotropic antiferromagnet at strong coupling. In panel (b) we compare the total energies across the 
    $\lambda_\mathrm{so}=0$ line to published data for Hartree-Fock,
    DMFT-NRG and AFQMC solutions \cite{Racz2020}.}
    \label{fig:KMenergy}
\end{figure}
